---
title: "V18 OpenPBTA Histologies File QC"
output: html_notebook    
toc: TRUE
toc_float: TRUE
editor_options: 
  chunk_output_type: inline
---
```{r global_options, include=F}
knitr::opts_chunk$set(error = F, echo = T, warning = T, message = T)
```


# Load packages
```{r}
suppressMessages(library(emo))
```

# Directories and Files
## Directories
```{r}
# Data directory
new_data_dir <- file.path("~/Documents/GitHub/jharenza.github.io/openPBTA-notebooks/V18-histologies-QC/data/")
pbta_data_dir <- file.path("~/OpenPBTA-analysis/data/release-v17-20200908/")
```

## Input Filepaths
```{r}
pbta_hist_file <- file.path(pbta_data_dir, "pbta-histologies.tsv")
hgg_hist_file <- file.path(new_data_dir, "pbta-histologies.tsv")
v18_hist_file <- file.path(new_data_dir, "Base Histologies Files Review - All.tsv")
```
## Read in files
```{r}
old_hist <- read.delim(pbta_hist_file, sep = "\t", header = T)
hgg_hist <- read.delim(hgg_hist_file, sep = "\t", header = T)
full_hist <- read.delim(v18_hist_file, sep = "\t", header = T)
```
## Subset new file for only those sampleIDs required
```{r}
new_hist <- full_hist[full_hist$Kids_First_Biospecimen_ID %in% hgg_hist$Kids_First_Biospecimen_ID,]
```

## First, assess dimensions whether new column names match the old
#### `r emo::ji("x")` Krutika needs to update column names: `os_days` -> `OS_days`, `os_status` -> `OS_days`, and `glioma_brain_region` -> `CNS_region`, add `pathology_free_text_diagnosis`, and Bailey update `cancer_predisposition` to `cancer_predispositions`

```{r}
dim(hgg_hist)
dim(new_hist)
setdiff(names(hgg_hist), names(new_hist))
setdiff(names(new_hist), names(hgg_hist))
```
```{r}
# I will update those for now so some of the below matching will work
colnames(hgg_hist)[colnames(hgg_hist) == "glioma_brain_region"] <- "CNS_region"
colnames(hgg_hist)[colnames(hgg_hist) == "os_days"] <- "OS_days"
colnames(hgg_hist)[colnames(hgg_hist) == "os_status"] <- "OS_status"
hgg_hist$pathology_free_text_diagnosis <- NA
colnames(new_hist)[colnames(new_hist) == "cancer_predisposition"] <- "cancer_predispositions"
```
## Check levels of histology columns
#### These levels look as expected `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(new_hist$pathology_diagnosis))
as.data.frame(table(new_hist$integrated_diagnosis))
as.data.frame(table(new_hist$broad_histology))
as.data.frame(table(new_hist$short_histology))
```
## Check tables from v17 -> v18 for fields not expected to change (quick overview)
#### Experimental strategy `r emo::ji("exclamation")`
```{r}
table(hgg_hist$experimental_strategy)
table(new_hist$experimental_strategy)
```
#### Extra RNA-Seq specimen in hgg `r emo::ji("exclamation")`
```{r}
hgg_rna <- subset(hgg_hist, experimental_strategy == "RNA-Seq")
new_rna <- subset(new_hist, experimental_strategy == "RNA-Seq")
setdiff(hgg_rna$Kids_First_Biospecimen_ID, new_rna$Kids_First_Biospecimen_ID) #no diff
hgg_rna[duplicated(hgg_rna$Kids_First_Biospecimen_ID),] #BS_NNPEC7W1
#temp remove row 242
hgg_hist <- hgg_hist[-242,]

```



#### Sample Type `r emo::ji("white_check_mark")`
```{r}
table(hgg_hist$sample_type)
table(new_hist$sample_type)
```
#### Composition `r emo::ji("x")`
There seem to be two blood samples that are now not reported in Bailey's table
```{r}
as.data.frame(table(hgg_hist$composition))
as.data.frame(table(new_hist$composition))
```
#### Broad composition `r emo::ji("x")`
Looks like this logic was missed
```{r}
as.data.frame(table(hgg_hist$broad_composition))
as.data.frame(table(new_hist$broad_composition))
```
#### Germline sex estimate - NA `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(hgg_hist$germline_sex_estimate))
as.data.frame(table(new_hist$germline_sex_estimate))
```
#### RNA library `r emo::ji("x")`
"Stranded"" needs case harmonization - should be lowercase "stranded" and polyA should be "poly-A"
```{r}
as.data.frame(table(hgg_hist$RNA_library))
as.data.frame(table(new_hist$RNA_library))
```
#### Cohort `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(hgg_hist$cohort))
as.data.frame(table(new_hist$cohort))
```
#### Sample and aliquot IDs `r emo::ji("white_check_mark")`
```{r}
setdiff(hgg_hist[, c("Kids_First_Biospecimen_ID", "sample_id", "aliquot_id", "cohort")],
        new_hist[, c("Kids_First_Biospecimen_ID", "sample_id", "aliquot_id", "cohort")])
setdiff(new_hist[, c("Kids_First_Biospecimen_ID", "sample_id", "aliquot_id", "cohort")],
        hgg_hist[, c("Kids_First_Biospecimen_ID", "sample_id", "aliquot_id", "cohort")])
```
#### Sequencing Center `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(hgg_hist$seq_center))
as.data.frame(table(new_hist$seq_center))
setdiff(hgg_hist[, c("Kids_First_Biospecimen_ID", "seq_center")],
        new_hist[, c("Kids_First_Biospecimen_ID", "seq_center")])
setdiff(new_hist[, c("Kids_First_Biospecimen_ID", "seq_center")],
        hgg_hist[, c("Kids_First_Biospecimen_ID", "seq_center")])
```
#### Tumor/Normal Fraction and Ploidy - NA for now
```{r}
setdiff(hgg_hist[, c("Kids_First_Biospecimen_ID", "normal_fraction", "tumor_fraction", "tumor_ploidy")],
        new_hist[, c("Kids_First_Biospecimen_ID", "normal_fraction", "tumor_fraction", "tumor_ploidy")])
setdiff(new_hist[, c("Kids_First_Biospecimen_ID", "normal_fraction", "tumor_fraction", "tumor_ploidy")],
        hgg_hist[, c("Kids_First_Biospecimen_ID", "normal_fraction", "tumor_fraction", "tumor_ploidy")]) 
```
#### Tumor Descriptor
```{r}
as.data.frame(table(hgg_hist$tumor_descriptor))
sum(as.data.frame(table(hgg_hist$tumor_descriptor))$Freq)
as.data.frame(table(new_hist$tumor_descriptor))
sum(as.data.frame(table(new_hist$tumor_descriptor))$Freq)

#CBTTC sample labeled as Initial CNS Tumor;Progressive and PNOC003 sample as Relapse were changed with last DS update before V17. Krutika will update to use the v17 release file
```
## Which are the samples that have had Tumor Descriptor changes?
#### 9 samples, mostly CBTTC, some PNOC003 `r emo::ji("exclamation")`
```{r}
setdiff(new_hist[,c("Kids_First_Biospecimen_ID", "tumor_descriptor", "cohort")], hgg_hist[,c("Kids_First_Biospecimen_ID", "tumor_descriptor", "cohort")])
```
## Did new subtypes get captured?
#### TBD `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(hgg_hist$molecular_subtype))
sum(as.data.frame(table(hgg_hist$molecular_subtype))$Freq)
as.data.frame(table(new_hist$molecular_subtype))
sum(as.data.frame(table(new_hist$molecular_subtype))$Freq)
```
## How many biospecimen pathology diagnoses were changed? `r emo::ji("white_check_mark")`
#### 7
```{r}
setdiff(new_hist[,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "cohort")], hgg_hist[,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "cohort")])
# Add to df
path_diff <- as.data.frame(setdiff(new_hist[,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "cohort")], hgg_hist[,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "cohort")]))
# How many per cohort?
table(path_diff$cohort)
```
#### Are the `integrated_diagnosis`, `broad_histology`, and `short_histology` NA for these samples? Yes `r emo::ji("white_check_mark")`
```{r}
new_hist[new_hist$Kids_First_Biospecimen_ID %in% path_diff$Kids_First_Biospecimen_ID,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "integrated_diagnosis", "broad_histology", "short_histology", "cohort")]
```


















