---
title: "V18 OpenPBTA Histologies File QC"
output: html_notebook    
toc: TRUE
toc_float: TRUE
editor_options: 
  chunk_output_type: inline
---
```{r global_options, include=F}
knitr::opts_chunk$set(error = F, echo = T, warning = T, message = T)
```


# Load packages
```{r}
suppressMessages(library(emo))
suppressMessages(library(tidyverse))

```

# Directories and Files
## Directories
```{r}
# Data directory
new_data_dir <- file.path("~/Documents/GitHub/jharenza.github.io/openPBTA-notebooks/V18-histologies-QC/data/")
pbta_data_dir <- file.path("~/Documents/GitHub/jharenza/OpenPBTA-analysis/data/release-v17-20200908")
```

## Input Filepaths
```{r}
pbta_hist_file <- file.path(pbta_data_dir, "pbta-histologies.tsv")
hgg_hist_file <- file.path("~/Documents/GitHub/d3b-pnoc003-omics/data/pbta-histologies.tsv")
v18_hist_file <- file.path(new_data_dir, "Base Histologies Files Review 20201116.tsv")
```
## Read in files
```{r}
old_hist <- read_tsv(pbta_hist_file, guess_max = 3000)
hgg_hist <- read_tsv(hgg_hist_file, guess_max = 3000)
full_hist <- read_tsv(v18_hist_file, guess_max = 10000)
```
## Subset new file for only those sampleIDs required
```{r}
new_ids <- read_tsv(file.path(new_data_dir, "2020-11-16-new-rna-seq.txt"), col_names = F) %>%
  pull(X1)
all_ids <- c(new_ids, old_hist$Kids_First_Biospecimen_ID)
new_hist <- full_hist %>%
  filter(Kids_First_Biospecimen_ID %in% all_ids)
```

## First, assess dimensions whether new column names match the old `r emo::ji("white_check_mark")`
```{r}
dim(old_hist)
dim(new_hist)
setdiff(names(new_hist), names(old_hist))
setdiff(names(old_hist), names(new_hist))
```

## Check levels of histology columns
#### These levels look as expected `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(new_hist$pathology_diagnosis))
as.data.frame(table(new_hist$integrated_diagnosis))
as.data.frame(table(new_hist$broad_histology))
as.data.frame(table(new_hist$short_histology))
```
## Check tables from v17 -> v18 for fields expected to change (8 more RNA-Seq)
#### Experimental strategy `r emo::ji("white_check_mark")`
```{r}
table(old_hist$experimental_strategy)
table(new_hist$experimental_strategy)
```

#### Sample Type `r emo::ji("white_check_mark")`
```{r}
table(old_hist$sample_type)
table(new_hist$sample_type)
```
#### CNS_region `r emo::ji("white_check_mark")`
# few updates
```{r}
table(old_hist$CNS_region)
table(new_hist$CNS_region)
```

#### Composition `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(old_hist$composition))
as.data.frame(table(new_hist$composition))
```
#### Broad composition `r emo::ji("white_check_mark")`
Removing in v18
```{r}
as.data.frame(table(old_hist$broad_composition))
as.data.frame(table(new_hist$broad_composition))
new_hist <- new_hist %>%
  select(-broad_composition)
```
#### Germline sex estimate - NA `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(old_hist$germline_sex_estimate))
as.data.frame(table(new_hist$germline_sex_estimate))
```
#### RNA library `r emo::ji("x")`
"Stranded" needs case harmonization - should be lowercase "stranded" and new samples are missing the library
```{r}
as.data.frame(table(old_hist$RNA_library))
as.data.frame(table(new_hist$RNA_library))

new_hist %>%
  filter(Kids_First_Biospecimen_ID %in% new_ids) %>%
  select(Kids_First_Biospecimen_ID, RNA_library)

# recode stranded library and add to samples missing it
new_hist <- new_hist %>%
  mutate(
    RNA_library = 
      case_when(RNA_library == "Stranded" | Kids_First_Biospecimen_ID %in% new_ids ~ "stranded", 
                TRUE ~ RNA_library))
```
#### Cohort `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(old_hist$cohort))
as.data.frame(table(new_hist$cohort))
```
#### Sample and aliquot IDs `r emo::ji("white_check_mark")`
```{r}
old_hist <- old_hist %>%
  arrange(Kids_First_Biospecimen_ID)
new_hist <- new_hist%>%
  arrange(Kids_First_Biospecimen_ID)
setdiff(old_hist[,c("sample_id", "aliquot_id")], new_hist[,c("sample_id", "aliquot_id")])
```
#### Sequencing Center `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(old_hist$seq_center))
as.data.frame(table(new_hist$seq_center))
```
#### Tumor/Normal Fraction and Ploidy - NA for now

#### Tumor Descriptor `r emo::ji("x")`
3 new PNOC BS_IDs do not have clinical metadata matched
```{r}
as.data.frame(table(old_hist$tumor_descriptor))
sum(as.data.frame(table(old_hist$tumor_descriptor))$Freq)
as.data.frame(table(new_hist$tumor_descriptor))
sum(as.data.frame(table(new_hist$tumor_descriptor))$Freq)
```

```{r}
missing_td <- new_hist %>%
  filter(tumor_descriptor == "Not Reported") %>%
  pull(Kids_First_Biospecimen_ID) %>%
  intersect(new_ids)

#Looks like primary site (Pons/Brainstem) and tumor descriptor (Initial CNS Tumor) is missing for these 3 samples - add back
new_hist <- new_hist %>%
  mutate(
    tumor_descriptor = 
      case_when(Kids_First_Biospecimen_ID %in% missing_td ~ "Initial CNS Tumor", 
                TRUE ~ tumor_descriptor),
    primary_site = 
      case_when(Kids_First_Biospecimen_ID %in% missing_td ~ "Pons/Brainstem",
                TRUE ~ primary_site))
```

## Did new subtypes get captured?

## How many biospecimen pathology diagnoses were changed? `r emo::ji("white_check_mark")`
#### 
```{r}
# Not sure why this is happening
setdiff(new_hist[,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "cohort")], old_hist[,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "cohort")])
```
#### Are the `integrated_diagnosis`, `broad_histology`, and `short_histology` NA for these samples? Yes `r emo::ji("white_check_mark")`
```{r}
#new_hist[new_hist$Kids_First_Biospecimen_ID %in% path_diff$Kids_First_Biospecimen_ID,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "integrated_diagnosis", "broad_histology", "short_histology", "cohort")]
```

# Remove BS_XM1AHBDJ for mapping rate < 60%
```{r}
#new_hist <- new_hist %>%
 # filter(Kids_First_Biospecimen_ID != "BS_XM1AHBDJ")
```

# Add back BIXU fields from V17
```{r}
names(old_hist)

# keeping int dx, broad/short hist for now just so CI does not break
old_fields <- old_hist %>%
  select(Kids_First_Biospecimen_ID, integrated_diagnosis, broad_histology, short_histology, germline_sex_estimate, normal_fraction, tumor_fraction, tumor_ploidy)

#remove those columns from V18 and replace
updated_columns <- new_hist %>%
  select(-colnames(old_fields[,2:ncol(old_fields)]))

# merge
final_hist <- updated_columns %>%
  left_join(old_fields, "Kids_First_Biospecimen_ID")
```

# Write new file
```{r}
write.table(final_hist, "~/Documents/GitHub/jharenza.github.io/openPBTA-notebooks/V18-histologies-QC/output/pbta-histologies.tsv", sep = "\t", quote = F, col.names = T, row.names = F)
```