---
title: "V17 OpenPBTA Histologies File QC"
output: html_notebook    
toc: TRUE
toc_float: TRUE
editor_options: 
  chunk_output_type: inline
---
```{r global_options, include=F}
knitr::opts_chunk$set(error = F, echo = T, warning = T, message = T)
```


# Load packages
```{r}
suppressMessages(library(emo))
```

# Directories and Files
## Directories
```{r}
# Data directory
new_data_dir <- file.path("~/Documents/GitHub/jharenza.github.io/openPBTA-notebooks/V17-histologies-QC/data/")
pbta_data_dir <- file.path("~/OpenPBTA-analysis/data/release-v16-20200320/")
```

## Input Filepaths
```{r}
pbta_hist_file <- file.path(pbta_data_dir, "pbta-histologies.tsv")
v17_hist_file <- file.path(new_data_dir, "Open PBTA v17 Candidate - v2.tsv")
```
## Read in files
```{r}
v16_hist <- read.delim(pbta_hist_file, sep = "\t", header = T)
v17_hist <- read.delim(v17_hist_file, sep = "\t", header = T)
```
## First, assess dimensions whether new column names match the old
#### New columns: `pathology_free_text_diagnosis` and `cohort_participant_id` and `glioma_brain_region` -> `CNS_region` `r emo::ji("white_check_mark")`
```{r}
dim(v16_hist)
dim(v17_hist)
setdiff(names(v16_hist), names(v17_hist))
setdiff(names(v17_hist), names(v16_hist))
```
## Check levels of histology columns
#### These levels look as expected `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(v17_hist$pathology_diagnosis))
as.data.frame(table(v17_hist$integrated_diagnosis))
as.data.frame(table(v17_hist$broad_histology))
as.data.frame(table(v17_hist$short_histology))
```
## Check tables from v16 -> v17 for fields not expected to change (quick overview)
#### Experimental strategy `r emo::ji("white_check_mark")`
```{r}
table(v16_hist$experimental_strategy)
table(v17_hist$experimental_strategy)
```
#### Sample Type `r emo::ji("white_check_mark")`
```{r}
table(v16_hist$sample_type)
table(v17_hist$sample_type)
```
#### Composition `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(v16_hist$composition))
as.data.frame(table(v17_hist$composition))
```
#### Broad composition `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(v16_hist$broad_composition))
as.data.frame(table(v17_hist$broad_composition))
```
#### Germline sex estimate `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(v16_hist$germline_sex_estimate))
as.data.frame(table(v17_hist$germline_sex_estimate))
```
#### RNA library `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(v16_hist$RNA_library))
as.data.frame(table(v17_hist$RNA_library))
```
#### Cohort `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(v16_hist$cohort))
as.data.frame(table(v17_hist$cohort))
```
#### Sample and aliquot IDs `r emo::ji("x")` - Aliquot ID for PNOC003 `BS_MVYA262V` sample `7316-3213` was changed from `A05265` to `A05233`. Is this correct? `r emo::ji("exclamation")`
```{r}
setdiff(v16_hist[, c("Kids_First_Biospecimen_ID", "sample_id", "aliquot_id", "cohort")],
        v17_hist[, c("Kids_First_Biospecimen_ID", "sample_id", "aliquot_id", "cohort")])
setdiff(v17_hist[, c("Kids_First_Biospecimen_ID", "sample_id", "aliquot_id", "cohort")],
        v16_hist[, c("Kids_First_Biospecimen_ID", "sample_id", "aliquot_id", "cohort")])
```
#### Sequencing Center - looks like `BS_QB84TBA3` and `BS_S2TA8R29` previously NA were updated to `NantOmics`. Is this correct? `r emo::ji("exclamation")`
```{r}
as.data.frame(table(v16_hist$seq_center))
as.data.frame(table(v17_hist$seq_center))
setdiff(v16_hist[, c("Kids_First_Biospecimen_ID", "seq_center")],
        v17_hist[, c("Kids_First_Biospecimen_ID", "seq_center")])
setdiff(v17_hist[, c("Kids_First_Biospecimen_ID", "seq_center")],
        v16_hist[, c("Kids_First_Biospecimen_ID", "seq_center")])
```
#### Tumor/Normal Fraction and Ploidy `r emo::ji("white_check_mark")`
```{r}
setdiff(v16_hist[, c("Kids_First_Biospecimen_ID", "normal_fraction", "tumor_fraction", "tumor_ploidy")],
        v17_hist[, c("Kids_First_Biospecimen_ID", "normal_fraction", "tumor_fraction", "tumor_ploidy")])
setdiff(v17_hist[, c("Kids_First_Biospecimen_ID", "normal_fraction", "tumor_fraction", "tumor_ploidy")],
        v16_hist[, c("Kids_First_Biospecimen_ID", "normal_fraction", "tumor_fraction", "tumor_ploidy")]) 
```
#### Tumor Descriptor `r emo::ji("x")`
```{r}
as.data.frame(table(v16_hist$tumor_descriptor))
sum(as.data.frame(table(v16_hist$tumor_descriptor))$Freq)
as.data.frame(table(v17_hist$tumor_descriptor))
sum(as.data.frame(table(v17_hist$tumor_descriptor))$Freq)
```
## Which are the samples that have had Tumor Descriptor changes?
#### 105 samples, mostly CBTTC - Seems like a lot, but could be reasonable with the refresh. Is this correct? `r emo::ji("exclamation")`
```{r}
setdiff(v17_hist[,c("Kids_First_Biospecimen_ID", "tumor_descriptor", "cohort")], v16_hist[,c("Kids_First_Biospecimen_ID", "tumor_descriptor", "cohort")])
```
## Did new subtypes get captured?
#### Went from 493 to 1015 subtypes, `BRAF` subtypes look to be updated, LGG subtypes added. `r emo::ji("white_check_mark")`
```{r}
as.data.frame(table(v16_hist$molecular_subtype))
sum(as.data.frame(table(v16_hist$molecular_subtype))$Freq)
as.data.frame(table(v17_hist$molecular_subtype))
sum(as.data.frame(table(v17_hist$molecular_subtype))$Freq)
```
## How many biospecimen pathology diagnoses were changed?
#### 33
```{r}
setdiff(v17_hist[,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "cohort")], v16_hist[,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "cohort")])
# Add to df
path_diff <- as.data.frame(setdiff(v17_hist[,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "cohort")], v16_hist[,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "cohort")]))
# How many per cohort?
table(path_diff$cohort)
```
#### Are the `integrated_diagnosis`, `broad_histology`, and `short_histology` NA for these samples? Looks like yes for CBTTC, but no for PNOC0003 `r emo::ji("exclamation")`
```{r}
v17_hist[v17_hist$Kids_First_Biospecimen_ID %in% path_diff$Kids_First_Biospecimen_ID,c("Kids_First_Biospecimen_ID", "pathology_diagnosis", "integrated_diagnosis", "broad_histology", "short_histology", "cohort")]
```
#### Note: For consistency of the process, we could make these NA for the PNOC003 samples for which `pathology_diagnosis` changed, however, the argument against this could be that we know all PNOC003 samples are DMGs and thus would not have `integrated_diagnosis`, `broad_histology`, and `short_histology` changes, so keeping these values and subtyping them from those values would be fine.

















