---
title: "Add plot labels to pbta-histologies.tsv file"
output: html_notebook    
toc: TRUE
toc_float: TRUE
editor_options: 
  chunk_output_type: inline
---

## Read in clinical data.

```{r}
clin <- read.delim("~/OpenPBTA-analysis/data/release-v15-20200228/pbta-histologies.tsv", header = T, sep = "\t")
```

## Add medulloblastoma labels.

```{r}
##If MB samples were not subtyped, retain only integrated diagnosis. If subtyped, concat with subtype
clin$label <- ifelse(clin$integrated_diagnosis == "Medulloblastoma" & !is.na(clin$molecular_subtype),
                       paste0(clin$integrated_diagnosis, "-", clin$molecular_subtype), 
                     ifelse(clin$integrated_diagnosis == "Medulloblastoma" & is.na(clin$molecular_subtype),
                            paste0(clin$integrated_diagnosis), paste0(clin$molecular_subtype)))
##check labels
as.data.frame(table(clin$label))
```

## Add ATRT labels.

```{r}
##Since ATRT has a broad_histology == Embryonal Tumor, we should specify the disease type in the label
clin$label <- unlist(ifelse(clin$short_histology == "ATRT", 
                            paste0(clin$short_histology), paste0(clin$label)))
##check labels
as.data.frame(table(clin$label))
```

## For "Other" tumors, use integrated diagnosis.
This results in a lot of histologies of small N - discuss what we would like to see here.

```{r}
clin$label <- unlist(ifelse(clin$short_histology == "Other", 
                            paste0(clin$integrated_diagnosis), paste0(clin$label)))
##check labels
as.data.frame(table(clin$label))
```

## Fix NAs - some NAs are converted to factors when the above is done (there may be a way to stop this during ifelse, but couldn't quickly find).

```{r}
clin$label <- ifelse(clin$label == "NA", NA, paste0(clin$label))
```

## For remaining samples, use broad_histology for missing labels.

```{r}
clin$label <- ifelse(is.na(clin$label), 
                     paste0(clin$short_histology), paste0(clin$label))
## Check all labels
as.data.frame(table(clin$label))
```

## A few notes:
- The 833 remaining NAs are the normal specimens.
- The CNS EFT-CIC will become CNS Ewings sarcoma.
- We have a standalone BRAF V600E subtype, but this is HGG only, and should be labeled as such.
- We probably want to label LGAT as BRAF fusion, BRAF V600E, or LGG BRAF WT.
